---
title: "Electricity Demand Forecast"
author: "Ananya Aggarwal_Jisup Kwak"
date: "2025-03-28"
output: html_document
---

## Setting R code chunk options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE)

```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(readxl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(tidyr)
library(dplyr)
library(imputeTS)
#New package for M9 to assist with tables
#install.packages("kableExtra")
library(kableExtra)
library(zoo)
library(writexl)

```

## Importing data

Hourly Electricity Demand, Temperature, Humidity from 01-01-2005 to 12-31-2010.

```{r}
# Organizing data of Electricity Demand
original_hourly_demand <- read_excel(path = "../Data/load.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)
original_hourly_demand <- original_hourly_demand[ , -1]


hourly_demand <- original_hourly_demand %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = "hour",
    values_to = "demand"
  )

hourly_demand <- hourly_demand %>%
  mutate(hour = as.integer(sub("h", "", hour)) - 1)

hourly_demand <- hourly_demand %>%
  mutate(
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, demand) %>%
  arrange(datetime)


ggplot(hourly_demand, aes(x = datetime, y = demand)) +
  geom_line(color = "steelblue") +
  labs(
    title = "Hourly Electricity Demand",
    x = "Datetime",
    y = "Demand (MW)"
  ) +
  theme_minimal()



```
# Cleaning data

```{r}
# Check NAs
summary(hourly_demand)
msts_demand_original <- msts(hourly_demand$demand,
                    seasonal.periods =c(24,168,8766),
                    start=c(2005,01,01))


# Check missing months
# 1. Ensure the 'datetime' column is in POSIXct format
hourly_demand$datetime <- as.POSIXct(hourly_demand$datetime)

# 2. Create a full sequence of hourly timestamps from the minimum to the maximum datetime
full_time <- seq(from = min(hourly_demand$datetime, na.rm = TRUE),
                 to   = max(hourly_demand$datetime, na.rm = TRUE),
                 by   = "hour")

# 3. Identify missing timestamps by comparing the full sequence with the existing datetimes
missing_times <- setdiff(full_time, hourly_demand$datetime)

# 4. Display the missing timestamps
missing_times

hourly_demand$datetime[which(hourly_demand$demand == 0)]
hourly_demand$datetime[which(is.na(hourly_demand$demand))]
# Replace 0 values in the demand column with NA
hourly_demand$demand[hourly_demand$demand < 1500] <- NA


# msts time series
msts_demand_nas <- msts(hourly_demand$demand,
                    seasonal.periods =c(24,168,8766),
                    start=c(2005,01,01))
head(msts_demand_nas)
autoplot(msts_demand_nas)

# datetime Zoo time series
zoo_demand <- zoo(hourly_demand$demand, order.by = hourly_demand$datetime)
head(zoo_demand)
tail(zoo_demand)

autoplot(zoo_demand) + 
  ggtitle("Original Hourly Demand Time Series")


# TS clean
# Convert to msts to allow seasonal-aware cleaning (e.g., daily and weekly patterns)

head(msts_demand_nas)
msts_demand_clean <- na.interp(msts_demand_nas)
summary(msts_demand_nas)
summary(msts_demand_clean)

# Convert cleaned series back to zoo format using original datetime index
zoo_demand_clean <- zoo(msts_demand_clean, order.by = index(zoo_demand))

index(zoo_demand_clean)[which(zoo_demand_clean == 0)]


# Plot the cleaned time series
autoplot(msts_demand_original, series = "Original") +
  autolayer(msts_demand_clean, series = "Cleaned") +
  labs(title = "Original vs Cleaned Time Series",
       x = "Datetime", y = "Demand") +
  scale_color_manual(values = c("Original" = "red", "Cleaned" = "blue")) +
  theme_minimal()



# Step 1: Create a dataframe with datetime and cleaned hourly demand
df_hourly <- data.frame(
  datetime = hourly_demand$datetime,
  demand_clean = as.numeric(msts_demand_clean)
)

# Step 2: Extract date
df_hourly <- df_hourly %>%
  mutate(date = as.Date(datetime))

# Step 3: Group by date and calculate daily average
df_daily_avg <- df_hourly %>%
  group_by(date) %>%
  summarise(daily_avg_demand = mean(demand_clean, na.rm = TRUE))

# Step 4: View result
head(df_daily_avg)

msts_daily_demand <- msts(df_daily_avg$daily_avg_demand,
                      start = c(2005, 1),
                      seasonal.periods = c(7, 365.25))


```

## Analyzing data

2005: 365days 2006: 365days 2007: 365days 2008: 366days 2009: 365days 2010: 365days

```{r}
# Hourly Electricity Demand

# ACF, PACF
par(mfrow=c(1,2))
ACF_Plot <- Acf(msts_demand_clean, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(msts_demand_clean, lag = 40)

```

# Creating a subset

```{r}
#create a subset for training purpose
n_for = 17530

msts_demand_train <- subset(msts_demand_clean,
  end = length(msts_demand_clean)-n_for)

#create a subset for testing purpose
msts_demand_test <- subset(msts_demand_clean,
                           start = length(msts_demand_clean)-n_for)


autoplot(msts_demand_train)
autoplot(msts_demand_test)


```

# Forecasting Hourly Electricity Demand

# Model 1: STL + ETS(ts object)

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(msts_demand_train, h=n_for)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(msts_demand) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Hourly Electricity Demand")

plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(ETS_fit$residuals), main = "ACF of Residuals")

names(ETS_fit)
length(ETS_fit$mean)

accuracy(ETS_fit$mean, msts_demand_test)

```

# Forecasting Hourly Electricity Demand
# model 1-1
```{r}

# Step 1: Fit ARIMA model with Fourier terms as external regressors
ARIMA_Fourier_fit <- auto.arima(msts_demand_train, seasonal = FALSE,
                                lambda=0,
                               xreg =
                                 fourier(msts_demand_train,
                                         K = c(2, 2, 1))
)


# Step 2: Forecast using the fitted model
ARIMA_Fourier_fc <- forecast(ARIMA_Fourier_fit, xreg =
                                 fourier(msts_demand_train,
                                         K = c(2, 2, 1),
                                         h = n_for), h=n_for)

# Step 3: Overlay forecast on full data
autoplot(msts_demand) +
  autolayer(ARIMA_Fourier_fc$mean, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Hourly Electricity Demand")

autoplot(ARIMA_Fourier_fc$mean)
# Step 8: Plot residuals
plot(ARIMA_Fourier_fc$residuals, main = "Residuals: ARIMA + Fourier")

# Step 9: Residual diagnostics
checkresiduals(ARIMA_Fourier_fc)
acf(na.omit(ARIMA_Fourier_fc$residuals), main = "ACF of Residuals")

# Step 10: Evaluate forecast accuracy
accuracy(ARIMA_Fourier_fc$mean, msts_demand_test)


```

# model 1-2: fitting
```{r}

# Step 1: Fit ARIMA model with Fourier terms as external regressors
ARIMA_Fourier_fit_2 <- auto.arima(msts_demand_train, seasonal = FALSE,
                                lambda=0,
                               xreg =
                                 fourier(msts_demand_train,
                                         K = c(4, 6, 12))
)


# Step 2: Forecast using the fitted model
ARIMA_Fourier_fc_2 <- forecast(ARIMA_Fourier_fit_2, xreg =
                                 fourier(msts_demand_train,
                                         K = c(4, 6, 12),
                                         h = n_for), h=n_for)

# Step 3: Overlay forecast on full data
autoplot(msts_demand) +
  autolayer(ARIMA_Fourier_fc_2$mean, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Hourly Electricity Demand")

autoplot(ARIMA_Fourier_fc_2$mean)
# Step 8: Plot residuals
plot(ARIMA_Fourier_fc_2$residuals, main = "Residuals: ARIMA + Fourier")

# Step 9: Residual diagnostics
checkresiduals(ARIMA_Fourier_fc_2)
acf(na.omit(ARIMA_Fourier_fc_2$residuals), main = "ACF of Residuals")

# Step 10: Evaluate forecast accuracy
accuracy(ARIMA_Fourier_fc_2$mean, msts_demand_test)


```
# model 1-3: forecasting
```{r}

# Step 1: Fit ARIMA model with Fourier terms as external regressors
# ARIMA_Fourier_fit_3 <- auto.arima(msts_demand_clean, 
#                                  seasonal = FALSE,
#                                  lambda=0,
#                                  xreg = fourier(msts_demand_clean,
#                                         K = c(4, 6, 12))
#)


# Step 2: Forecast using the fitted model
ARIMA_Fourier_fc_3 <- forecast(ARIMA_Fourier_fit_3, xreg =
                                 fourier(msts_demand_clean,
                                         K = c(4, 6, 12)), h=1416)

# Step 3: Overlay forecast on full data
autoplot(msts_demand) +
  autolayer(ARIMA_Fourier_fc_3$mean, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Hourly Electricity Demand")

autoplot(ARIMA_Fourier_fc_3$mean)
# Step 8: Plot residuals
plot(ARIMA_Fourier_fit_3$residuals, main = "Residuals: ARIMA + Fourier")

# Step 9: Residual diagnostics
checkresiduals(ARIMA_Fourier_fc_3)
acf(na.omit(ARIMA_Fourier_fc_3$residuals), main = "ACF of Residuals")




# 1. Create hourly datetime starting from "2011-01-01 01:00:00"
forecast_times <- seq(from = as.POSIXct("2011-01-01 01:00:00"),
                      by = "hour",
                      length.out = length(ARIMA_Fourier_fc_3$mean))

# 2. Create dataframe with forecast values and datetime
forecast_df <- data.frame(
  datetime = forecast_times,
  hourly_forecast = as.numeric(ARIMA_Fourier_fc_3$mean)
)

# 3. Extract date only
forecast_df <- forecast_df %>%
  mutate(date = as.Date(datetime))

# 4. Compute daily average demand
daily_avg_df <- forecast_df %>%
  group_by(date) %>%
  summarise(daily_avg_demand = mean(hourly_forecast, na.rm = TRUE))

# 5. Save to Excel
write_xlsx(daily_avg_df, "daily_forecast_average.xlsx")

```
# model 1-4: forecasting
```{r}


# Generate Fourier terms
xreg_train <- fourier(msts_daily_demand, K = c(2, 5))

# Fit ARIMA model with Fourier terms
ARIMA_Fourier_fit_4 <- auto.arima(msts_daily_demand,
                                  seasonal = FALSE,
                                  lambda = 0,
                                  xreg = xreg_train)

# Step 2: Forecast using the fitted model (for daily data)
n_forecast_days <- 59  # or however many days you want to forecast

# Generate future Fourier terms
xreg_future <- fourier(msts_daily_demand, K = c(2, 5), h = n_forecast_days)

# Forecast
ARIMA_Fourier_fc_4 <- forecast(ARIMA_Fourier_fit_4,
                               xreg = xreg_future,
                               h = n_forecast_days)

# Step 3: Plot full forecast result
autoplot(ARIMA_Fourier_fc_4) +
  ylab("Daily Electricity Demand") +
  ggtitle("Forecast: ARIMA + Fourier (Daily)")

# Step 4: Plot residuals
plot(ARIMA_Fourier_fit_4$residuals, main = "Residuals: ARIMA + Fourier")

# Step 5: Residual diagnostics
checkresiduals(ARIMA_Fourier_fit_4)
acf(na.omit(ARIMA_Fourier_fit_4$residuals), main = "ACF of Residuals")


# 1. Create forecast date sequence starting from the day after the last observation
start_date <- as.Date(tail(df_daily_avg$date, 1)) + 1

# 2. Generate corresponding date vector
forecast_dates <- seq.Date(from = start_date,
                           by = "day",
                           length.out = n_forecast_days)

# 3. Create result dataframe
daily_forecast_df <- data.frame(
  date = forecast_dates,
  forecast_demand = as.numeric(ARIMA_Fourier_fc_4$mean)
)

# 4. Save to Excel
write_xlsx(daily_forecast_df, "daily_forecast_average.xlsx")


```


# Forecasting with Neural Network methods
Model 2-1
```{r}

# Step 1: Generate Fourier terms for training set
xreg_train <- fourier(msts_daily_demand, K = c(2, 5))

# Step 2: Fit Neural Network Time Series Model with Fourier terms
NN_fit <- nnetar(msts_daily_demand,
                 p = 1,
                 P = 0,
                 xreg = xreg_train,
                 lambda = 0)  # Box-Cox transform if needed


# Forecast horizon
n_forecast_days <- 59

# Generate future Fourier terms
xreg_future <- fourier(msts_daily_demand, K = c(2, 5), h = n_forecast_days)

# Forecast using the neural network model
NN_fc <- forecast(NN_fit,
                  xreg = xreg_future,
                  h = n_forecast_days)
autoplot(NN_fc) +
  ylab("Daily Electricity Demand") +
  ggtitle("Forecast: Neural Network + Fourier (Daily)")
plot(NN_fit$residuals, main = "Residuals: Neural Network + Fourier")

acf(na.omit(NN_fit$residuals), main = "ACF of Residuals")
# Create date sequence for forecast horizon
start_date <- as.Date(tail(df_daily_avg$date, 1)) + 1
forecast_dates <- seq.Date(from = start_date,
                           by = "day",
                           length.out = n_forecast_days)

# Create output dataframe
daily_forecast_df <- data.frame(
  date = forecast_dates,
  forecast_demand = as.numeric(NN_fc$mean)
)

# Save to Excel
write_xlsx(daily_forecast_df, "daily_forecast_nnetar_fourier.xlsx")

```
Model 2-2
```{r}

# Step 1: Generate Fourier terms for training set
xreg_train <- fourier(msts_daily_demand, K = c(2, 4))

# Step 2: Fit Neural Network Time Series Model with Fourier terms
NN_fit <- nnetar(msts_daily_demand,
                 p = 1,
                 P = 0,
                 xreg = xreg_train,
                 lambda = 0)  # Box-Cox transform if needed


# Forecast horizon
n_forecast_days <- 59

# Generate future Fourier terms
xreg_future <- fourier(msts_daily_demand, K = c(2, 4), h = n_forecast_days)

# Forecast using the neural network model
NN_fc <- forecast(NN_fit,
                  xreg = xreg_future,
                  h = n_forecast_days)
autoplot(NN_fc) +
  ylab("Daily Electricity Demand") +
  ggtitle("Forecast: Neural Network + Fourier (Daily)")
plot(NN_fit$residuals, main = "Residuals: Neural Network + Fourier")

acf(na.omit(NN_fit$residuals), main = "ACF of Residuals")
# Create date sequence for forecast horizon
start_date <- as.Date(tail(df_daily_avg$date, 1)) + 1
forecast_dates <- seq.Date(from = start_date,
                           by = "day",
                           length.out = n_forecast_days)

# Create output dataframe
daily_forecast_df <- data.frame(
  date = forecast_dates,
  forecast_demand = as.numeric(NN_fc$mean)
)

# Save to Excel
write_xlsx(daily_forecast_df, "daily_forecast_nnetar_fourier_2-2.xlsx")

```
