---
title: "Electricity Demand Forecast"
author: "Ananya Aggarwal_Jisup Kwak"
date: "2025-03-28"
output: html_document
---

## Setting R code chunk options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE)
knitr::opts_knit$set(root.dir = "~/R/Time Series Analysis/TSA_Final/KwakAggarwal_ENV797_TSA_ForecastCompetition_S25")
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(readxl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(tidyr)
library(dplyr)
library(imputeTS)
#New package for M9 to assist with tables
#install.packages("kableExtra")
library(kableExtra)

```

## Importing data

Hourly Electricity Demand, Temperature, Humidity from 01-01-2005 to 12-31-2010.

```{r}
# Organizing data of Electricity Demand
original_hourly_demand <- read_excel(path = "./Data/load.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)
original_hourly_demand <- original_hourly_demand[ , -1]


hourly_demand <- original_hourly_demand %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = "hour",
    values_to = "demand"
  )

hourly_demand <- hourly_demand %>%
  mutate(hour = as.integer(sub("h", "", hour)) - 1)

hourly_demand <- hourly_demand %>%
  mutate(
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, demand) %>%
  arrange(datetime)


# Organizing data of Temperature
original_hourly_temp <- read_excel(path = "./Data/temperature.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_temp <- original_hourly_temp %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("t_ws"))


# Organizing data of Relative Humidity
original_hourly_humid <- read_excel(path = "./Data/relative_humidity.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_humid <- original_hourly_humid %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("rh_ws"))


```

## Analyzing data
2005: 365days
2006: 365days
2007: 365days
2008: 366days
2009: 365days
2010: 365days

```{r}
# Hourly Electricity Demand
summary(hourly_demand)
na_demand <- sum(is.na(hourly_demand$demand))
which(is.na(hourly_demand$demand))
mean_demand <- mean(hourly_demand$demand, na.rm = TRUE)

# Initial Plots
ggplot(hourly_demand, aes(x = datetime, y = demand)) +
  geom_line(color = "darkblue") +
  geom_hline(yintercept = mean_demand, color = "red", size = 0.8) +
  labs(title = "Hourly Electricity Demand",
       subtitle = paste("Mean Demand =", round(mean_demand, 2), "MW"),
       x = "Datetime", y = "Demand (MW)") +
  theme_minimal()

# ACF, PACF
par(mfrow=c(1,2))
ACF_Plot <- Acf(hourly_demand$demand, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(hourly_demand$demand, lag = 40)

# Transforming the series into time series object.
ts_hourly_demand <- msts(hourly_demand$demand,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))
ts_hourly_demand <- na_interpolation(ts_hourly_demand, option = "linear")
head(ts_hourly_demand)
summary(ts_hourly_demand)

# Decomposing
# seasonal-24
# seasonal-168
# seasonal-8766
ts_hourly_demand %>% mstl() %>%
autoplot()





```
# Creating a subset


```{r}
#create a subset for training purpose
n_for = 8765
ts_hourly_demand_train <- subset(ts_hourly_demand,
  end = length(ts_hourly_demand)-n_for)
#create a subset for testing purpose
ts_hourly_demand_test <- subset(ts_hourly_demand,
start = length(ts_hourly_demand)-n_for)
autoplot(ts_hourly_demand_train)
autoplot(ts_hourly_demand_test)

```

# Forecasting Hourly Electricity Demand
# Model 1: STL + ETS

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(ts_hourly_demand_train, h=8766)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(ts_hourly_demand) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# 실제와 예측값을 하나의 데이터프레임으로 결합
df_compare <- data.frame(
  Time = time(ts_hourly_demand_test),
  Actual = as.numeric(ts_hourly_demand_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# 예측구간에 대한 비교 그래프
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()

df_compare$Residuals <- df_compare$Actual - df_compare$Forecast
ggplot(df_compare, aes(x = Time, y = Residuals)) +
  geom_line(color = "darkblue") +
  ylab("Residuals") +
  ggtitle("Residual Plot: Actual - Forecast (STL + ETS)") +
  theme_minimal()
ggplot(df_compare, aes(x = Residuals)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  ggtitle("Histogram of Residuals") +
  xlab("Residuals") +
  theme_minimal()
plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

names(ETS_fit)
ETS_fit$fitted

accuracy_result <- accuracy(ETS_fit$mean, ts_hourly_demand_test)
print(accuracy_result)
```


