---
title: "Electricity Demand Forecast"
author: "Ananya Aggarwal_Jisup Kwak"
date: "2025-03-28"
output: html_document
---

## Setting R code chunk options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE)

```

## Loading packages and initializing

```{r package, message=FALSE}
library(readxl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(tidyr)
library(dplyr)
library(imputeTS)
library(kableExtra)
library(writexl)
library(openxlsx)
```

## Importing data
Hourly Electricity Demand, Temperature, Humidity from 01-01-2005 to 12-31-2010.

```{r}
# Load hourly electricity demand data
original_hourly_demand <- read_excel(path = "../Data/load.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)
original_hourly_demand <- original_hourly_demand[ , -1]

# Reshape demand data from wide to long format
hourly_demand <- original_hourly_demand %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = "hour",
    values_to = "demand"
  )

hourly_demand <- hourly_demand %>%
  mutate(hour = as.integer(sub("h", "", hour)) - 1)

hourly_demand <- hourly_demand %>%
  mutate(
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, demand) %>%
  arrange(datetime)

# Plot hourly electricity demand
ggplot(hourly_demand, aes(x = datetime, y = demand)) +
  geom_line(color = "steelblue") +
  labs(
    title = "Hourly Electricity Demand",
    x = "Datetime",
    y = "Demand (MW)"
  ) +
  theme_minimal()

# Load hourly temperature data
original_hourly_temp <- read_excel(path = "../Data/temperature.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_temp <- original_hourly_temp %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("t_ws"))


hourly_temp_long <- hourly_temp %>%
  pivot_longer(
    cols = starts_with("t_ws"),
    names_to = "station",
    values_to = "temperature"
  )


ggplot(hourly_temp_long, aes(x = datetime, y = temperature, color = station)) +
  geom_line(alpha = 0.6) +
  labs(
    title = "Hourly Temperature at 28 Weather Stations",
    x = "Datetime",
    y = "Temperature (°F)",
    color = "Station"
  ) +
  theme_minimal()


# Hourly average temp
hourly_temp_average <- data.frame(
  datetime = hourly_temp$datetime,
  mean_temp = rowMeans(hourly_temp[,2:29], na.rm = TRUE)
)


ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "tomato") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()

# Load relative humidity data
original_hourly_humid <- read_excel(path = "../Data/relative_humidity.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_humid <- original_hourly_humid %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("rh_ws"))
summary(hourly_humid)

# Hourly average humidity
hourly_humid_average <- data.frame(
  datetime = hourly_humid$datetime,
  mean_humid = rowMeans(hourly_humid[,2:29], na.rm = TRUE)
)

ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "orange") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()


```
# Cleaning data

```{r}
# Convert to time series object with daily seasonality (24*31 for monthly)
summary(hourly_demand)
ts_demand <- ts(hourly_demand$demand, frequency = 24*31)
summary(ts_demand)

# Interpolate missing values
ts_demand <- na.interp(ts_demand)
summary(ts_demand)

# STL decomposition for trend, seasonality, and remainder
decomp <- stl(ts_demand, s.window = "periodic")

trend <- decomp$time.series[, "trend"]
seasonal <- decomp$time.series[, "seasonal"]
remainder <- decomp$time.series[, "remainder"]

# Identify outliers: demand <= 1000 with large residuals
outlier_index <- which(ts_demand <= 1000)


# Replace outliers with NA and interpolate
ts_demand_fixed <- ts_demand
ts_demand_fixed[outlier_index] <- NA
ts_demand_clean <- na.interp(ts_demand_fixed)


df_plot <- data.frame(
  time = time(ts_demand),
  original = as.numeric(ts_demand),
  cleaned = as.numeric(ts_demand_clean)
)

plot(ts_demand_clean)
ggplot(df_plot, aes(x = time)) +
  geom_line(aes(y = original), color = "gray", alpha = 0.5) +
  geom_line(aes(y = cleaned), color = "blue") +
  geom_point(data = df_plot[outlier_index, ], aes(y = original), color = "red", size = 1.5) +
  ggtitle("STL-based Outlier Replacement (<=1000 & large residual)") +
  ylab("Electricity Demand") +
  theme_minimal()
summary(ts_demand_clean)
sum(ts_demand_clean <= 1000, na.rm = TRUE)


## Time Series Transformation(MSTS)
msts_hourly_demand <- msts(hourly_demand$demand,
seasonal.periods =c(24,168,8760),
start=c(2005,01,01))
msts_hourly_temp_average <- msts(hourly_temp_average$mean_temp,
seasonal.periods =c(24,168,8760),
start=c(2005,01,01))
msts_hourly_humid_average <- msts(hourly_humid_average$mean_humid,
seasonal.periods =c(24,168,8760),
start=c(2005,01,01))



# Replace low demand values (<= 1000) with NA and interpolate
msts_demand_na <- msts_hourly_demand
msts_demand_na[msts_demand_na <= 1000] <- NA
msts_demand_clean <- na.interp(msts_demand_na)

# Plot cleaned vs original demand data
df_plot <- data.frame(
  time = time(msts_hourly_demand),
  original = as.numeric(msts_hourly_demand),
  cleaned = as.numeric(msts_demand_clean)
)

ggplot(df_plot, aes(x = time)) +
  geom_line(aes(y = original), color = "gray", alpha = 0.5) +
  geom_line(aes(y = cleaned), color = "blue") +
  geom_point(data = df_plot[df_plot$original <= 1000, ], aes(y = original), color = "red", size = 0.7) +
  ggtitle("STL-based Cleaning: Values <= 1000 Replaced") +
  ylab("Electricity Demand") +
  theme_minimal()

plot(msts_demand_clean)


```

## Analyzing data

2005: 365days 2006: 365days 2007: 365days 2008: 366days 2009: 365days 2010: 365days

```{r}

# ACF, PACF
par(mfrow=c(1,2))
ACF_Plot <- Acf(ts_demand_clean, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(ts_demand_clean, lag = 40)


```

## Splitting Data into Training and Test Sets

```{r}
#create a subset for training purpose
n_for = 8760
summary(ts_demand_clean)

ts_demand_clean_train <- subset(ts_demand_clean,
  end = length(ts_demand_clean)-n_for)

#create a subset for testing purpose
ts_demand_clean_test <- subset(ts_demand_clean,
start = length(ts_demand_clean)-n_for)

autoplot(ts_demand_clean_train)
autoplot(ts_demand_clean_test)


#create a subset for training purpose
n_for = 8760
msts_demand_clean_train <- subset(msts_demand_clean,
  end = length(msts_demand_clean)-n_for)
msts_hourly_temp_avaerage_train <- subset(msts_hourly_temp_average,
  end = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_train <- subset(msts_hourly_humid_average,
  end = length(msts_hourly_humid_average)-n_for)


#create a subset for testing purpose
msts_demand_clean_test <- subset(msts_demand_clean,
start = length(msts_demand_clean)-n_for)
msts_hourly_temp_average_test <- subset(msts_hourly_temp_average,
start = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_test <- subset(msts_hourly_humid_average,
start = length(msts_hourly_humid_average)-n_for)



```

# Forecasting Hourly Electricity Demand

# Model 1: STL + ETS(ts object) test
This model is not good for forecasting. The reason is that ts can has only one time period.

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(ts_demand_clean_train, h=8761)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot forecast with full data
autoplot(ts_demand_clean) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")

# Evaluate forecast accuracy
df_compare <- data.frame(
  Time = time(ts_demand_clean_test),
  Actual = as.numeric(ts_demand_clean_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# Plot actual vs forecast
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()



checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

# accuracy
ETS_forecast_ts <- ts(ETS_fit$mean, start = start(ts_demand_clean_test), frequency = frequency(ts_demand_clean_test))

accuracy_result <- accuracy(ETS_forecast_ts, ts_demand_clean_test)

print(accuracy_result)
```



# Model 2: STL + ETS(msts object)
msts can have multiple time periods.
```{r}
# Model 2: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(msts_demand_clean_train, h=8761)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(msts_demand_clean) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# for comparison
df_compare <- data.frame(
  Time = time(msts_demand_clean_test),
  Actual = as.numeric(msts_demand_clean_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# plot
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()


checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

#accuracy_result <- accuracy(ETS_fit$mean, msts_demand_clean_test)
#print(accuracy_result)

```
# Model 2: STL + ETS(MSTS) forecast
```{r}
ETS_fit <- stlf(msts_demand_clean, h = 8760)
forecast_8760 <- ETS_fit$mean[1:8760]
time_seq <- seq(from = time(ETS_fit$mean)[1], by = 1/24, length.out = 8760)
df_forecast_export <- data.frame(Time = time_seq, Forecast = as.numeric(forecast_8760))
write_xlsx(df_forecast_export, path = "stl_ets_forecast_8760.xlsx")

```


# Forecasting Hourly Electricity Demand

# Model 2: ARIMA+Fourier(test)
```{r}

# Fit ARIMA + Fourier model on training set
ARIMA_Four_fit <- auto.arima(msts_demand_clean_train,
                             seasonal=FALSE, lambda=0,
                             xreg=fourier(msts_demand_clean_train,
                                          K=c(4,6,8)),
                             stepwise = TRUE,
                             approximation = TRUE
)

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(msts_demand_clean_train,
                                        K=c(4,6,8),
                                        h=8671),
                           h=8761)
#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Active Power")

# Step 5: 예측 정확도 확인
acc <- accuracy(ARIMA_Four_for$mean, msts_demand_clean_test)
print(acc)

# Step 6: 시각화
autoplot(msts_demand_clean_test, series = "Actual") +
  autolayer(ARIMA_Four_for$mean, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("ARIMA + Fourier Forecast vs Actual") +
  scale_color_manual(values = c("Actual" = "black", "ARIMA + Fourier" = "blue")) +
  theme_minimal()


```
# Model 2: ARIMA+Fourier(forecast)

```{r}
# Fit ARIMA + Fourier model on training set
ARIMA_Four_fit_12m <- auto.arima(msts_demand_clean,
                             seasonal=FALSE, lambda=0,
                             xreg=fourier(msts_demand_clean,
                                          K=c(4,6,12)),
                             stepwise = TRUE,
                             approximation = TRUE
)
#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for_12m <- forecast(ARIMA_Four_fit_12m,
                           xreg=fourier(msts_demand_clean,
                                        K=c(4,6,12),
                                        h=59),
                           h=59)

# Step 8: 예측 결과를 데이터프레임으로 변환
forecast_df <- data.frame(
  Time = time(ARIMA_Four_for_12m$mean),
  Forecast = as.numeric(ARIMA_Four_for_12m$mean)
)
plot(ARIMA_Four_for_12m$mean)
# Step 9: 엑셀로 저장
write.xlsx(forecast_df, file = "ARIMA_Fourier_12month_forecast.xlsx")

```


# TBATS

```{r}
# TBATS can take time to fit
#TBATS_fit <- tbats(msts_demand_clean_train)
#TBATS_for <- forecast(TBATS_fit, h=8761)
#Plot foresting results
#autoplot(TBATS_for) +
#ylab("Active Power")

#Plot model + observed data
#autoplot(msts_demand_clean) +
#autolayer(TBATS_for, series="TBATS",PI=FALSE)+
#ylab("Active Power")

# Step 4: TBATS 모델로 향후 8760시간 (1년) 예측
TBATS_fit_59 <- tbats(msts_demand_clean)
TBATS_future <- forecast(TBATS_fit_8760, h = 59)

autoplot(msts_demand_clean) +
autolayer(TBATS_future, series="TBATS",PI=FALSE)+
ylab("Active Power")

# Step 5: 예측 결과를 데이터프레임으로 정리
TBATS_future_df <- data.frame(
  Time = time(TBATS_future$mean),
  Forecast = as.numeric(TBATS_future$mean)
)

# Step 6: 엑셀로 저장 (openxlsx 패키지 사용)
# install.packages("openxlsx")  # 필요 시 설치


write.xlsx(TBATS_future_df, file = "TBATS_8760hour_forecast.xlsx")

# 완료 메시지 출력
cat("TBATS 기반 8760시간 예측이 'TBATS_8760hour_forecast.xlsx'로 저장되었습니다.\n")




```

# Forecasting with Neural Network methods

```{r}
NN_fit <- nnetar(msts_demand_clean,
                  p=1,
                  P=0,
                  xreg=fourier(msts_demand_clean, K=c(1,2,3)))

NN_for <- forecast(NN_fit, h=59,xreg=fourier(msts_demand_clean,
                    K=c(1,2,3),h=59))
#Plot foresting results
autoplot(NN_for) +
ylab("Active Power")

# Plot model + observed data
autoplot(msts_demand_clean) +
  autolayer(NN_for$mean, series = "Neural Network", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Set)") +
  theme_minimal()


# 예측 결과만 시각화
autoplot(NN_for) +
  ylab("Electricity Demand") +
  ggtitle("NNAR + Fourier Forecast")

# 실제 시계열과 예측치 함께 시각화
autoplot(msts_demand_clean) +
  autolayer(NN_for$mean, series = "Neural Network (NNAR + Fourier)", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Period)") +
  theme_minimal()

# Step 5: 예측 결과를 데이터프레임으로 정리
NN_for_df <- data.frame(
  Time = time(NN_for$mean),
  Forecast = as.numeric(NN_for$mean)
)
plot(NN_for$mean)
write.xlsx(NN_for_df, file = "TBATS_8760hour_forecast.xlsx")

```

NN_forecast


```{r}
NN_fit <- nnetar(msts_demand_clean,
                  p=1,
                  P=0,
                  xreg=fourier(msts_demand_clean_train, K=c(5,3,2)))

NN_for <- forecast(NN_fit, h=8760,xreg=fourier(msts_demand_clean_train,
                    K=c(5,3,2),h=8760))
#Plot foresting results
autoplot(NN_for) +
ylab("Active Power")

# Plot model + observed data
autoplot(msts_demand_clean_test) +
  autolayer(NN_for$mean, series = "Neural Network", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Set)") +
  theme_minimal()


# 예측 결과만 시각화
autoplot(NN_for) +
  ylab("Electricity Demand") +
  ggtitle("NNAR + Fourier Forecast")

# 실제 시계열과 예측치 함께 시각화
autoplot(msts_demand_clean_test) +
  autolayer(NN_for$mean, series = "Neural Network (NNAR + Fourier)", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Period)") +
  theme_minimal()


```



```{r}

```

