---
title: "Electricity Demand Forecast"
author: "Ananya Aggarwal_Jisup Kwak"
date: "2025-03-28"
output: html_document
---

## Setting R code chunk options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE)

```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(readxl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(tidyr)
library(dplyr)
library(imputeTS)
#New package for M9 to assist with tables
#install.packages("kableExtra")
library(kableExtra)

```

## Importing data

Hourly Electricity Demand, Temperature, Humidity from 01-01-2005 to 12-31-2010.

```{r}
# Organizing data of Electricity Demand
original_hourly_demand <- read_excel(path = "../Data/load.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)
original_hourly_demand <- original_hourly_demand[ , -1]


hourly_demand <- original_hourly_demand %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = "hour",
    values_to = "demand"
  )

hourly_demand <- hourly_demand %>%
  mutate(hour = as.integer(sub("h", "", hour)) - 1)

hourly_demand <- hourly_demand %>%
  mutate(
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, demand) %>%
  arrange(datetime)


ggplot(hourly_demand, aes(x = datetime, y = demand)) +
  geom_line(color = "steelblue") +
  labs(
    title = "Hourly Electricity Demand",
    x = "Datetime",
    y = "Demand (MW)"
  ) +
  theme_minimal()

# Organizing data of Temperature
original_hourly_temp <- read_excel(path = "../Data/temperature.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_temp <- original_hourly_temp %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("t_ws"))

# pivot_longer transform
hourly_temp_long <- hourly_temp %>%
  pivot_longer(
    cols = starts_with("t_ws"),
    names_to = "station",
    values_to = "temperature"
  )

# 시각화
ggplot(hourly_temp_long, aes(x = datetime, y = temperature, color = station)) +
  geom_line(alpha = 0.6) +
  labs(
    title = "Hourly Temperature at 28 Weather Stations",
    x = "Datetime",
    y = "Temperature (°F)",
    color = "Station"
  ) +
  theme_minimal()


# Hourly average temp
hourly_temp_average <- data.frame(
  datetime = hourly_temp$datetime,
  mean_temp = rowMeans(hourly_temp[,2:29], na.rm = TRUE)
)

# 시각화
ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "tomato") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()


# Organizing data of Relative Humidity
original_hourly_humid <- read_excel(path = "../Data/relative_humidity.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_humid <- original_hourly_humid %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("rh_ws"))
summary(hourly_humid)
# Hourly average temp
hourly_humid_average <- data.frame(
  datetime = hourly_humid$datetime,
  mean_humid = rowMeans(hourly_humid[,2:29], na.rm = TRUE)
)

# 시각화
ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "orange") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()


```
# Cleaning data

```{r}
summary(hourly_demand)
ts_demand <- ts(hourly_demand$demand, frequency = 24*31)
summary(ts_demand)
ts_demand <- na.interp(ts_demand)
summary(ts_demand)

# Step 2: STL 분해
decomp <- stl(ts_demand, s.window = "periodic")

# Step 3: 구성요소 확인
trend <- decomp$time.series[, "trend"]
seasonal <- decomp$time.series[, "seasonal"]
remainder <- decomp$time.series[, "remainder"]

# Step 4: 이상치 기준 정의: demand <= 1000 이면서 잔차가 매우 큰 경우
outlier_index <- which(ts_demand <= 1000)


# Step 5: 이상치 대체 - NA로 만들고 보간
ts_demand_fixed <- ts_demand
ts_demand_fixed[outlier_index] <- NA
ts_demand_clean <- na.interp(ts_demand_fixed)

# Step 6: 시각화
df_plot <- data.frame(
  time = time(ts_demand),
  original = as.numeric(ts_demand),
  cleaned = as.numeric(ts_demand_clean)
)

plot(ts_demand_clean)
ggplot(df_plot, aes(x = time)) +
  geom_line(aes(y = original), color = "gray", alpha = 0.5) +
  geom_line(aes(y = cleaned), color = "blue") +
  geom_point(data = df_plot[outlier_index, ], aes(y = original), color = "red", size = 1.5) +
  ggtitle("STL-based Outlier Replacement (<=1000 & large residual)") +
  ylab("Electricity Demand") +
  theme_minimal()
summary(ts_demand_clean)
sum(ts_demand_clean <= 1000, na.rm = TRUE)


# Transforming the series into time series object.
msts_hourly_demand <- msts(hourly_demand$demand,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))
msts_hourly_temp_average <- msts(hourly_temp_average$mean_temp,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))
msts_hourly_humid_average <- msts(hourly_humid_average$mean_humid,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))


head(msts_hourly_demand_ip)
summary(msts_hourly_demand_ip)

# Step 2: 1000 이하 값 NA로 처리
msts_demand_na <- msts_hourly_demand
msts_demand_na[msts_demand_na <= 1000] <- NA

# Step 3: NA 보간 (STL 기반으로 추세 + 계절 반영)
msts_demand_clean <- na.interp(msts_demand_na)

# Step 4: 시각화 (원본 vs 수정)
df_plot <- data.frame(
  time = time(msts_hourly_demand),
  original = as.numeric(msts_hourly_demand),
  cleaned = as.numeric(msts_demand_clean)
)

ggplot(df_plot, aes(x = time)) +
  geom_line(aes(y = original), color = "gray", alpha = 0.5) +
  geom_line(aes(y = cleaned), color = "blue") +
  geom_point(data = df_plot[df_plot$original <= 1000, ], aes(y = original), color = "red", size = 0.7) +
  ggtitle("STL-based Cleaning: Values <= 1000 Replaced") +
  ylab("Electricity Demand") +
  theme_minimal()

plot(msts_demand_clean)


```

## Analyzing data

2005: 365days 2006: 365days 2007: 365days 2008: 366days 2009: 365days 2010: 365days

```{r}

# ACF, PACF
par(mfrow=c(1,2))
ACF_Plot <- Acf(ts_demand_clean, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(ts_demand_clean, lag = 40)


```

# Creating a subset

```{r}
#create a subset for training purpose
n_for = 8760
summary(ts_demand_clean)

ts_demand_clean_train <- subset(ts_demand_clean,
  end = length(ts_demand_clean)-n_for)

#create a subset for testing purpose
ts_demand_clean_test <- subset(ts_demand_clean,
start = length(ts_demand_clean)-n_for)

autoplot(ts_demand_clean_train)
autoplot(ts_demand_clean_test)


#create a subset for training purpose
n_for = 8760
msts_demand_clean_train <- subset(msts_demand_clean,
  end = length(msts_demand_clean)-n_for)
msts_hourly_temp_avaerage_train <- subset(msts_hourly_temp_average,
  end = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_train <- subset(msts_hourly_humid_average,
  end = length(msts_hourly_humid_average)-n_for)


#create a subset for testing purpose
msts_demand_clean_test <- subset(msts_demand_clean,
start = length(msts_demand_clean)-n_for)
msts_hourly_temp_average_test <- subset(msts_hourly_temp_average,
start = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_test <- subset(msts_hourly_humid_average,
start = length(msts_hourly_humid_average)-n_for)



```

# Forecasting Hourly Electricity Demand

# Model 1: STL + ETS(ts object)

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(ts_demand_clean_train, h=8761)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(ts_demand_clean) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# 실제와 예측값을 하나의 데이터프레임으로 결합
df_compare <- data.frame(
  Time = time(ts_demand_clean_test),
  Actual = as.numeric(ts_demand_clean_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# 예측구간에 대한 비교 그래프
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()

df_compare$Residuals <- df_compare$Actual - df_compare$Forecast
ggplot(df_compare, aes(x = Time, y = Residuals)) +
  geom_line(color = "darkblue") +
  ylab("Residuals") +
  ggtitle("Residual Plot: Actual - Forecast (STL + ETS)") +
  theme_minimal()
ggplot(df_compare, aes(x = Residuals)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  ggtitle("Histogram of Residuals") +
  xlab("Residuals") +
  theme_minimal()
plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

names(ETS_fit)
ETS_fit$fitted

length(ETS_fit$mean)
length(ts_demand_clean_test)

# ts_hourly_demand_test와 같은 주기 및 시작 시점을 사용
ETS_forecast_ts <- ts(ETS_fit$mean, start = start(ts_demand_clean_test), frequency = frequency(ts_demand_clean_test))

# accuracy 계산
accuracy_result <- accuracy(ETS_forecast_ts, ts_demand_clean_test)

print(accuracy_result)
```
# Model 1: STL + ETS(msts object)

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(msts_demand_clean_train, h=8761)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(msts_demand_clean) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# 실제와 예측값을 하나의 데이터프레임으로 결합
df_compare <- data.frame(
  Time = time(msts_demand_clean_test),
  Actual = as.numeric(msts_demand_clean_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# 예측구간에 대한 비교 그래프
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()

df_compare$Residuals <- df_compare$Actual - df_compare$Forecast
ggplot(df_compare, aes(x = Time, y = Residuals)) +
  geom_line(color = "darkblue") +
  ylab("Residuals") +
  ggtitle("Residual Plot: Actual - Forecast (STL + ETS)") +
  theme_minimal()
ggplot(df_compare, aes(x = Residuals)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  ggtitle("Histogram of Residuals") +
  xlab("Residuals") +
  theme_minimal()
plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

names(ETS_fit)
ETS_fit$fitted

accuracy_result <- accuracy(ETS_fit$mean, msts_demand_clean_test)
print(accuracy_result)

```
# Forecasting Hourly Electricity Demand

# Model 2: ARIMA+Fourier
```{r}

# Step 1: Fourier terms 설정 (예: 월=744, 주=168, 일=24)
ARIMA_Four_fit <- auto.arima(msts_demand_clean_train,
                             seasonal=FALSE, lambda=0,
                             xreg=fourier(msts_demand_clean_train,
                                          K=c(2,3,3)),
                             stepwise = TRUE,
                             approximation = TRUE
)

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(msts_demand_clean_train,
                                        K=c(2,3,3),
                                        h=8671),
                           h=8761)
#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Active Power")

# Step 5: 예측 정확도 확인
acc <- accuracy(ARIMA_Four_for$mean, msts_demand_clean_test)
print(acc)

# Step 6: 시각화
autoplot(msts_demand_clean_test, series = "Actual") +
  autolayer(ARIMA_Four_for$mean, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("ARIMA + Fourier Forecast vs Actual") +
  scale_color_manual(values = c("Actual" = "black", "ARIMA + Fourier" = "blue")) +
  theme_minimal()


```

# TBATS

```{r}
# TBATS can take time to fit
TBATS_fit <- tbats(msts_demand_clean_train)
TBATS_for <- forecast(TBATS_fit, h=8761)
#Plot foresting results
autoplot(TBATS_for) +
ylab("Active Power")

#Plot model + observed data
autoplot(msts_demand_clean) +
autolayer(TBATS_for, series="TBATS",PI=FALSE)+
ylab("Active Power")
```

# Forecasting with Neural Network methods

```{r}
NN_fit <- nnetar(msts_demand_clean_train,
                  p=1,
                  P=0,
                  xreg=fourier(msts_demand_clean_train, K=c(5,3,2)))

NN_for <- forecast(NN_fit, h=8760,xreg=fourier(msts_demand_clean_train,
                    K=c(5,3,2),h=8760))
#Plot foresting results
autoplot(NN_for) +
ylab("Active Power")

# Plot model + observed data
autoplot(msts_demand_clean_test) +
  autolayer(NN_for$mean, series = "Neural Network", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Set)") +
  theme_minimal()


# 예측 결과만 시각화
autoplot(NN_for) +
  ylab("Electricity Demand") +
  ggtitle("NNAR + Fourier Forecast")

# 실제 시계열과 예측치 함께 시각화
autoplot(msts_demand_clean_test) +
  autolayer(NN_for$mean, series = "Neural Network (NNAR + Fourier)", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Period)") +
  theme_minimal()
```

