---
title: "Electricity Demand Forecast"
author: "Ananya Aggarwal_Jisup Kwak"
date: "2025-03-28"
output: html_document
---

## Setting R code chunk options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE)

```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(readxl)
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)
library(tidyr)
library(dplyr)
library(imputeTS)
#New package for M9 to assist with tables
#install.packages("kableExtra")
library(kableExtra)

```

## Importing data

Hourly Electricity Demand, Temperature, Humidity from 01-01-2005 to 12-31-2010.

```{r}
# Organizing data of Electricity Demand
original_hourly_demand <- read_excel(path = "../Data/load.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)
original_hourly_demand <- original_hourly_demand[ , -1]


hourly_demand <- original_hourly_demand %>%
  pivot_longer(
    cols = starts_with("h"),
    names_to = "hour",
    values_to = "demand"
  )

hourly_demand <- hourly_demand %>%
  mutate(hour = as.integer(sub("h", "", hour)) - 1)

hourly_demand <- hourly_demand %>%
  mutate(
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, demand) %>%
  arrange(datetime)


ggplot(hourly_demand, aes(x = datetime, y = demand)) +
  geom_line(color = "steelblue") +
  labs(
    title = "Hourly Electricity Demand",
    x = "Datetime",
    y = "Demand (MW)"
  ) +
  theme_minimal()

# Organizing data of Temperature
original_hourly_temp <- read_excel(path = "../Data/temperature.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_temp <- original_hourly_temp %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("t_ws"))

# pivot_longer transform
hourly_temp_long <- hourly_temp %>%
  pivot_longer(
    cols = starts_with("t_ws"),
    names_to = "station",
    values_to = "temperature"
  )

# 시각화
ggplot(hourly_temp_long, aes(x = datetime, y = temperature, color = station)) +
  geom_line(alpha = 0.6) +
  labs(
    title = "Hourly Temperature at 28 Weather Stations",
    x = "Datetime",
    y = "Temperature (°F)",
    color = "Station"
  ) +
  theme_minimal()


# Hourly average temp
hourly_temp_average <- data.frame(
  datetime = hourly_temp$datetime,
  mean_temp = rowMeans(hourly_temp[,2:29], na.rm = TRUE)
)

# 시각화
ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "tomato") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()


# Organizing data of Relative Humidity
original_hourly_humid <- read_excel(path = "../Data/relative_humidity.xlsx", skip = 0, sheet = "Sheet1", col_names = TRUE)

hourly_humid <- original_hourly_humid %>%
  mutate(
    hour = hr - 1,
    datetime = ymd(date) + hours(hour)
  ) %>%
  select(datetime, starts_with("rh_ws"))
summary(hourly_humid)
# Hourly average temp
hourly_humid_average <- data.frame(
  datetime = hourly_humid$datetime,
  mean_humid = rowMeans(hourly_humid[,2:29], na.rm = TRUE)
)

# 시각화
ggplot(hourly_temp_average, aes(x = datetime, y = mean_temp)) +
  geom_line(color = "orange") +
  labs(
    title = "Hourly Average Temperature (Across 28 Weather Stations)",
    x = "Datetime",
    y = "Average Temperature (°F)"
  ) +
  theme_minimal()


```
# Cleaning data

```{r}
summary(hourly_demand)
hourly_demand_fixed <- hourly_demand %>%
  mutate(demand = ifelse(demand <= 1000, 2331, demand))
summary(hourly_demand_fixed)
plot(hourly_demand_fixed$demand)
which(hourly_demand$demand == 0)
which(is.na(hourly_demand$demand))


hourly_demand_ip <- data.frame(
  datetime = hourly_demand_fixed$datetime,
  demand_ip = na_interpolation(hourly_demand_fixed$demand, option = "linear"))
mean_demand = mean(hourly_demand_ip$demand_ip)
summary(hourly_demand_ip)
which(hourly_demand_ip$demand == 0)
plot(hourly_demand_ip$demand)


summary(hourly_demand_ip)
which(hourly_demand_ip$demand == 0)
plot(hourly_demand_ip$demand)
```

## Analyzing data

2005: 365days 2006: 365days 2007: 365days 2008: 366days 2009: 365days 2010: 365days

```{r}
# Hourly Electricity Demand



# transforming Time Series
ts_hourly_demand_ip <- ts(hourly_demand_ip$demand_ip,start=c(2005,1),frequency=24*31) 
summary(ts_hourly_demand_ip)
ts_hourly_humid_average <- ts(hourly_humid_average$mean_humid, start=c(2005,1),frequency=24*31) 
summary(ts_hourly_humid_avaerage)
ts_hourly_temp_average <- ts(hourly_temp_average$mean_temp, start=c(2005,1),frequency=24*31) 
summary(ts_hourly_temp_average)


decom_ts_hourly_demand_ip <- decompose(ts_hourly_demand_ip)
plot(decom_ts_hourly_demand_ip)

# Initial Plots
ggplot(hourly_demand_ip, aes(x = datetime, y = demand_ip)) +
  geom_line(color = "darkblue") +
  geom_hline(yintercept = mean_demand, color = "red", size = 0.8) +
  labs(title = "Hourly Electricity Demand",
       subtitle = paste("Mean Demand =", round(mean_demand, 2), "MW"),
       x = "Datetime", y = "Demand (MW)") +
  theme_minimal()

# ACF, PACF
par(mfrow=c(1,2))
ACF_Plot <- Acf(hourly_demand_ip$demand_ip, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(hourly_demand_ip$demand_ip, lag = 40)

# Transforming the series into time series object.
msts_hourly_demand_ip <- msts(hourly_demand_ip$demand_ip,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))
msts_hourly_temp_average <- msts(hourly_temp_average$mean_temp,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))
msts_hourly_humid_average <- msts(hourly_humid_average$mean_humid,
seasonal.periods =c(24,168,8766),
start=c(2005,01,01))

head(msts_hourly_demand_ip)
summary(msts_hourly_demand_ip)

# Decomposing
# seasonal-24
# seasonal-168
# seasonal-8766

```

# Creating a subset

```{r}
#create a subset for training purpose
n_for = 8765
summary(ts_hourly_demand_ip)
summary(hourly_demand_ip)
summary(hourly_demand)

ts_hourly_demand_ip_train <- subset(ts_hourly_demand_ip,
  end = length(ts_hourly_demand_ip)-n_for)
ts_hourly_temp_avaerage_train <- subset(ts_hourly_temp_average,
  end = length(ts_hourly_temp_average)-n_for)
ts_hourly_humid_average_train <- subset(ts_hourly_humid_average,
  end = length(ts_hourly_humid_average)-n_for)


#create a subset for testing purpose
ts_hourly_demand_ip_test <- subset(ts_hourly_demand_ip,
start = length(ts_hourly_demand_ip)-n_for)
ts_hourly_temp_average_test <- subset(ts_hourly_temp_average,
start = length(ts_hourly_temp_average)-n_for)
ts_hourly_humid_average_test <- subset(ts_hourly_humid_average,
start = length(ts_hourly_humid_average)-n_for)



autoplot(ts_hourly_demand_ip_train)
autoplot(ts_hourly_demand_ip_test)
autoplot(ts_hourly_temp_avaerage_train)
autoplot(ts_hourly_temp_average_test)
autoplot(ts_hourly_humid_average_train)
autoplot(ts_hourly_humid_average_test)


#create a subset for training purpose
n_for = 8765
msts_hourly_demand_ip_train <- subset(msts_hourly_demand_ip,
  end = length(msts_hourly_demand_ip)-n_for)
msts_hourly_temp_avaerage_train <- subset(msts_hourly_temp_average,
  end = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_train <- subset(msts_hourly_humid_average,
  end = length(msts_hourly_humid_average)-n_for)


#create a subset for testing purpose
msts_hourly_demand_ip_test <- subset(msts_hourly_demand_ip,
start = length(msts_hourly_demand_ip)-n_for)
msts_hourly_temp_average_test <- subset(msts_hourly_temp_average,
start = length(msts_hourly_temp_average)-n_for)
msts_hourly_humid_average_test <- subset(msts_hourly_humid_average,
start = length(msts_hourly_humid_average)-n_for)




```

# Forecasting Hourly Electricity Demand

# Model 1: STL + ETS(ts object)

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(ts_hourly_demand_ip_train, h=8766)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(ts_hourly_demand_ip) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# 실제와 예측값을 하나의 데이터프레임으로 결합
df_compare <- data.frame(
  Time = time(ts_hourly_demand_ip_test),
  Actual = as.numeric(ts_hourly_demand_ip_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# 예측구간에 대한 비교 그래프
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()

df_compare$Residuals <- df_compare$Actual - df_compare$Forecast
ggplot(df_compare, aes(x = Time, y = Residuals)) +
  geom_line(color = "darkblue") +
  ylab("Residuals") +
  ggtitle("Residual Plot: Actual - Forecast (STL + ETS)") +
  theme_minimal()
ggplot(df_compare, aes(x = Residuals)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  ggtitle("Histogram of Residuals") +
  xlab("Residuals") +
  theme_minimal()
plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

names(ETS_fit)
ETS_fit$fitted

length(ETS_fit$mean)
length(ts_hourly_demand_test)

# ts_hourly_demand_test와 같은 주기 및 시작 시점을 사용
ETS_forecast_ts <- ts(ETS_fit$mean, start = start(ts_hourly_demand_test), frequency = frequency(ts_hourly_demand_test))

# accuracy 계산
accuracy_result <- accuracy(ETS_forecast_ts, ts_hourly_demand_test)

print(accuracy_result)
```
# Model 1: STL + ETS(msts object)

```{r}
# Model 1: STL + ETS
#Fit and forecast STL + ETS model to data
ETS_fit <- stlf(msts_hourly_demand_ip_train, h=8766)
#Plot foresting results
autoplot(ETS_fit) + ylab("Hourly Electricity Demand")
#Plot model + observed data
autoplot(msts_hourly_demand_ip) +
autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
ylab("Hourly Electricity Demand")


# 실제와 예측값을 하나의 데이터프레임으로 결합
df_compare <- data.frame(
  Time = time(msts_hourly_demand_ip_test),
  Actual = as.numeric(msts_hourly_demand_ip_test),
  Forecast = as.numeric(ETS_fit$mean)
)

# 예측구간에 대한 비교 그래프
ggplot(df_compare, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecast, color = "Forecast")) +
  ylab("Hourly Electricity Demand") +
  ggtitle("Forecast vs Actual (Forecast Horizon Only)") +
  scale_color_manual(name = "", values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()

df_compare$Residuals <- df_compare$Actual - df_compare$Forecast
ggplot(df_compare, aes(x = Time, y = Residuals)) +
  geom_line(color = "darkblue") +
  ylab("Residuals") +
  ggtitle("Residual Plot: Actual - Forecast (STL + ETS)") +
  theme_minimal()
ggplot(df_compare, aes(x = Residuals)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  ggtitle("Histogram of Residuals") +
  xlab("Residuals") +
  theme_minimal()
plot(ETS_fit$residuals)
checkresiduals(ETS_fit)
acf(na.omit(df_compare$Residuals), main = "ACF of Residuals")

names(ETS_fit)
ETS_fit$fitted

accuracy_result <- accuracy(ETS_fit$mean, msts_hourly_demand_ip_test)
print(accuracy_result)

```
# Forecasting Hourly Electricity Demand

# Model 2: ARIMA+Fourier
```{r}
anyNA(msts_hourly_demand_ip_train)       # TRUE면 NA 있음
any(is.nan(msts_hourly_demand_ip_train)) # NaN 있음?
any(is.infinite(msts_hourly_demand_ip_train)) # Inf 있음?
summary(msts_hourly_demand_ip_train)

# ARIMA + Fourier terms 모델 적합
ARIMA_Four_fit <- auto.arima(
  msts_hourly_demand_ip_train,
  seasonal = FALSE,
  lambda = 0,
  xreg = fourier(msts_hourly_demand_ip_train, K = c(2, 3, 4))
)

# 예측 수행 (테스트 구간 길이 기준)
ARIMA_Four_for <- forecast(
  ARIMA_Four_fit,
  xreg = fourier(msts_hourly_demand_ip_train, K = c(2, 3, 4), h = n_for),
  h = n_for
)

# 예측 결과 시각화
autoplot(ARIMA_Four_for) + ylab("Electricity Demand")

# 전체 시계열 + 예측치 시각화
autoplot(msts_hourly_demand_ip) +
  autolayer(ARIMA_Four_for, series = "ARIMA + Fourier", PI = FALSE) +
  ylab("Electricity Demand")


```

# SARIMA

```{r}
#SARIMA_autofit <- auto.arima(ts_hourly_demand_ip)
#checkresiduals(SARIMA_autofit)

```

# Forecasting with Neural Network methods

```{r}
#NN_fit <- nnetar(ts_hourly_demand_ip_train,p=1,P=1)
NN_fit <- nnetar(ts_hourly_demand_ip_train,
                  p=1,
                  P=0,
                  xreg=fourier(ts_hourly_demand_train, K=c(5,3,2)))

#NN_for <- forecast(NN_fit, h=8760)
NN_for <- forecast(NN_fit, h=8760,xreg=fourier(ts_hourly_demand_train,
                    K=c(5,3,2),h=8760))
#Plot foresting results
autoplot(NN_for) +
ylab("Active Power")

# Plot model + observed data
autoplot(ts_hourly_demand_ip_test) +
  autolayer(NN_for$mean, series = "Neural Network", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Set)") +
  theme_minimal()


# 예측 결과만 시각화
autoplot(NN_for) +
  ylab("Electricity Demand") +
  ggtitle("NNAR + Fourier Forecast")

# 실제 시계열과 예측치 함께 시각화
autoplot(ts_hourly_demand_ip_test) +
  autolayer(NN_for$mean, series = "Neural Network (NNAR + Fourier)", PI = FALSE) +
  ylab("Hourly Electricity Demand") +
  ggtitle("NNAR Forecast vs Actual (Test Period)") +
  theme_minimal()
```

```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
str(ts_hourly_demand_train)
attributes(ts_hourly_demand_train)
ts_hourly_demand_train_weekly <- ts(as.numeric(ts_hourly_demand_train),
                                    frequency = 168, start = c(2005, 1))
NN_fit <- nnetar(ts_hourly_demand_train_weekly,
                  p=1,
                  P=1,
                  lambda = NULL)
#NN_for <- forecast(NN_fit, h=365)
NN_for <- forecast(NN_fit, h=8760)
#Plot foresting results
autoplot(NN_for) +
ylab("Active Power")

#Plot model + observed data
autoplot(ts_hourly_demand) +
autolayer(NN_for, series="Neural Network",PI=FALSE)+
ylab("Hourly Electricity Demand")


```
